---
title: "Data_access_cleaning"
output: html_document
---

```{r}
#installing the relavent packages
library(tidyverse)

#install relavent packages to access aws
install.packages("aws.s3", repos = c("cloudyr" = "http://cloudyr.github.io/drat"))
library(aws.s3)

source('C:/Users/Sreekanth/Desktop/DSSG Project/key.txt')

Sys.setenv("AWS_ACCESS_KEY_ID" = access_key,
           "AWS_SECRET_ACCESS_KEY" = secret_key,
           "AWS_DEFAULT_REGION" = "us-west-2")

get_bucket('dssg2018')

data <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Choice_1314-1718.csv")

head(data)
```

```{r}
names(data)
```

```{r}
summary(data)
```



Some exploratory data analysis.
What are the unique values of the columns in the dataset that makes sense?
```{r}
#Unique school year
columns = c("SchoolYear","PrimaryHomeLanguage",  "Grade", "ChoiceSelection", "ChoiceStatus")

finding_unique = function(z, data) {
  unique(data[,z])
}

# Creating a list using lapply

unique_values_columns = sapply(columns, finding_unique, data = data)
unique_values_columns
```

What are the data types of each of the columns in the dataset?
```{r}
sapply(data, typeof)
```

Are there any missing values in the dataset?

```{r}
#Replacing all the missing values with NA
data = lapply(data, function(x) {x[x == " "] <- NA})
sapply(data, function(x) sum(is.na(x)))
```

****************************************Looking at the Attendence dataset********************************************

```{r}

#Skipping the first row inorder to g et the appropriate headers
DPS_Attendance <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_Attendance_1112-1718.csv", skip = 1, na.strings = " ") 

#Assigning appropriate column names
colnames(DPS_Attendance) = c("StudentNumber", "2011-2012", "2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018")
colnames(DPS_Attendance) <- make.names(colnames(DPS_Attendance))

#Replacing all the missing values with NA
DPS_Attendance = apply(DPS_Attendance, 2, function(x) gsub("^$|^ $", NA, x))

DPS_Attendance = as.data.frame(DPS_Attendance)
head(DPS_Attendance)
```

Checking if the student attendance data is distinct on the studentnumber level.
```{r}
nrow(DPS_Attendance)
length(unique(DPS_Attendance$StudentNumber))
length(which(is.na(DPS_Attendance$StudentNumber)))

```

The data is unique at the stuent number level and also no missing student numbers. 
```{r}
summary(DPS_Attendance)
```

We need discuss if the number of missing records in each of the years is a cause of concern. 

```{r}
nrow(DPS_Attendance[DPS_Attendance$na_count > 6,])
```
There are 90 student records that have attendance records missing or all the years. We might not necessarily need to have them in the dataset but we can look at relavent information for these students in other datasets and then take a call. 

Finding the minimum and maximum of attendence in each of the years and noting any discrepencies
```{r}

#Ignoring the studentnumbers
x = DPS_Attendance[,-1]

#Storing the minimum and maximum of each column as a dataframe
min_values = apply(x, 2, min, na.rm = TRUE)
min_values = as.data.frame(min_values)

max_values = apply(x, 2, max, na.rm = TRUE)
max_values = as.data.frame(max_values)

na_count <-apply(x, function(y) sum(length(which(is.na(y)))))
min_max_info = cbind(min_values, max_values)


```
As we can see academic years 2014-2015 and 2016-2017 have incorrect minimum values which needs to be addressed.




*******************************Looking at the rawdata/DPS_CMAS_1415-1617.csv dataset*****************************************

```{r}
#Skipping the first row inorder to g et the appropriate headers
data1 <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_CMAS_1415-1617.csv", na.strings = " ") 
data2 <- s3read_using(FUN = read.csv, object = "s3://dssg2018/rawdata/DPS_CMAS_1415-1617.csv", skip = 1, na.strings = " ")
#Assigning appropriate column names
colnames(DPS_Attendance) = c("StudentNumber", "2011-2012", "2012-2013", "2013-2014", "2014-2015", "2015-2016", "2016-2017", "2017-2018")
colnames(DPS_Attendance) <- make.names(colnames(DPS_Attendance))

#Replacing all the missing values with NA
DPS_Attendance = apply(DPS_Attendance, 2, function(x) gsub("^$|^ $", NA, x))

DPS_Attendance = as.data.frame(DPS_Attendance)
head(data)
```

```{r}
unique(data$X.6)
```

